---
title: "A2"
author: "Jenny Ly"
date: "16/03/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

#                      Truck
#Oil | 1         2      3       4        5
# 1  | 0.500   0.634   0.487   0.329    0.512
# 2  | 0.535   0.675   0.520   0.435    0.540
# 3  | 0.513   0.595   0.488   0.400    0.510
#--------------------------------------------------

```{r}
#Question 1
#Analyze the data form this experiement (alpha=0.05)
y <- c(0.500,0.634,0.487,0.329,0.512,
       0.535,0.675,0.520,0.435,0.540,
       0.513,0.595,0.488,0.400,0.510)
oil.level <- rep(1:3,each=5)
truck.level <- rep(1:5,3)

oil <- factor(oil.level) #treatment = 3
truck <- factor(truck.level) #block = 5

fit <- aov(y~oil+truck,contrasts = list(oil=contr.sum, truck=contr.sum) )
anova(fit)

#base on ANOVa table, at level 0.05, the treatment means different significantly

########################################################
#B-Use the Fisher LSD method 
#to make comparisions among the three lubricaitng oils 
#to determine specifically which oils differ in break-specific fuel consumption 

library("agricolae")

test <- LSD.test(fit,trt = "oil", p.adj = "none", group = FALSE)
test

#at level at 0.05, there's significant difference 
#between oil 1&2, and oil number 2&3 in terms of break-specific fuel consumption

#C.Analyze the residuals from this experiment 

# Standarized residuals and fitted values:
rcbd.r1<-rstandard(fit)
rcbd.fitted1<-fitted.values(fit)

# Normal Q-Q plot:
qqnorm(rcbd.r1,main="Normal Q-Q Plot")
qqline(rcbd.r1)

#normal Q-Q plot indicates an approximate straight line relationship 
#suggests that the error distribution is approximately normal

#Plot of residuals vs. fitted values:
plot(rcbd.fitted1,rcbd.r1,ylab="Standardized residuals", 
  xlab="Fitted values",main="Plot of residuals vs. fitted values")
abline(h=0)
#Comment: The plot of residuals vs fitted values verifies constant assumption

# Plot of residuals vs. treatment:
plot(oil.level,rcbd.r1,ylab="Standardized residuals", 
  xlab=" Lubricating Oil",main="Plot of residuals vs. Treatment")
abline(h=0)
#Comment: No indication of inequality of variance by treatment


# Plot of residuals vs. block:
plot(truck.level,rcbd.r1,ylab="Standardized residuals", 
  xlab="Truck Engines",main="Plot of residuals vs. Block")
abline(h=0)
#comment: Nothing unusual, all the blocks seem homogeneous 
```

```{r}
#Question 2
#a/ Analyze the data from this experiment, alpha = 0.05 
batchs <- factor(c(rep(1:5,5)))
day <- factor(c(rep(1:5, each=5)))
ingre <- factor(c("A","C","B","D","E",
           "B","E","A","C","D",
           "D","A","C","E","B",
           "C","D","E","B","A",
           "E","B","D","A","C"))
hour <- c(8,11,4,6,4,
          7,2,9,8,2,
          1,7,10,6,3,
          7,3,1,6,8,
          3,8,5,10,8)

myfit <- lm(hour~batchs+day+ingre)
anova(myfit)
#base on ANOVA table, at 95% CI treatment means differ significantly
#Differences between batches is not significant at level 0.05
#Also at level 0.05 no significant difference between days

```
#Suppose observation from batch 3 on day 4 is missing. Estimate missing value and perform analysis using this value

$p = 5$

$row.prime = 4+9+10+5 = 28$

$column.prime = 7+3+6+8=24$

$treament.prime = 4+2+6+3=15$

$grandtotal.prime=8+7+1+7+...+8 = 146$

$\hat{y}_{ijk}=\frac{5.(28+24+15)-2.146}{(5-2)(5-1)}=3.58=4$

```{r}
batch2 <- factor(c(rep(1:5,5)))
day2 <- factor(c(rep(1:5, each=5)))
ingre2 <- factor(c("A","C","B","D","E",
           "B","E","A","C","D",
           "D","A","C","E","B",
           "C","D","E","B","A",
           "E","B","D","A","C"))
hour2 <- c(8,11,4,6,4,
          7,2,9,8,2,
          1,7,10,6,3,
          7,3,4,6,8,
          3,8,5,10,8)

myfit2 <- lm(hour2~batch2+day2+ingre2)
anova(myfit2)

#at 0.05 level, there's still significant difference in treatment means
#batches and days effects are relatively small
```
```{r}
#Question 3 
#Analyze the data from this experiment (alpha = 0.05)
trt <- c(2,4,6,8,10,12,14)
k <- 3
bib <- design.bib(trt,k,r=NULL)
bib


```
$H_0:\tau_1=\tau_2=\tau_3=\tau_4=\tau_5=0$  $H_1:$ at least one $\tau_i \ne 0$

$SS_T=(114)^2+(120)^2+....+(127)^2-\frac{2715}{21}=2600.29$

$SS_{Blocks}={\frac13}[(381)^2+....+(378)^2]-\frac{{(2715)}^2}{21}=1114.29$

$Q_1=351-\frac13[381+0+0+0+381+0+378]=-29$

$Q_2=365-\frac13[381+402+0+0+0+372+0]=-20$

$Q_3=388-\frac13[0+402+366+0+0+0+378]=6$

$Q_4=25$
$Q_5=32$
$Q_6=-12$
$Q_7=-2$

$SS_{treatment}=\frac{(-29)^2+(-20)^2+(6^2)+25^2+32^2+(-12^2+(-2^2))}{1.7}=1317.43$

$SS_E=2600.29-1114.29-1317.43=168.57$

$MS_E=\frac{168.57}{8}=21.07$

$MS_{treatment}=\frac{1317.43}{6}=219.57$

$MS_{blocks}=\frac{1114.29}{6}=185.72$

$F_o=\frac{219.57}{21.07}=10.42$

$F_{(6,8)}=2.66833$

$F_0>F_{(6,8)}$

#reject hypothesis.At level 0.05 there's signifcant difference in treatments means 

#MS_blocks/MS_E does not close to 1, so blocking is efficient in this model

```{r}
#Question 4
#a. Analyze the data and draw conclusions. alpha = 0,05
p1.y <- c(90.4,90.2,90.7,90.6,90.2,90.4,
        90.1,90.3,90.5,90.6,89.9,90.1,
        90.5,90.7,90.8,90.9,90.4,90.1)

p1.A <- rep(c(150,160,170), each = 6)
p1.B <- rep(rep(c(200,215,230),each= 2),3)
p1.Af<-factor(p1.A)
p1.Bf<-factor(p1.B)


p1.fit1<-lm(p1.y~p1.Af*p1.Bf,
  contrasts=list(p1.Af=contr.sum,p1.Bf=contr.sum))

p1.anova<-anova(p1.fit1)
p1.anova
#From ANOVA table,the interaction of Temp and Press is insignificant at level 0.05
#Temperature and Pressure means are significant difference at level 0.05 
#other methods of comparison can be performed for further analysis 

#b.Draw appropriate residual plots and comment on the model's adequacy
p1.r1<-rstandard(p1.fit1)
p1.fitted1<-fitted.values(p1.fit1)

par(mfrow=c(2,2))

# Normal Q-Q plot
qqnorm(p1.r1,ylab="Standardized residuals",xlab="Theoretical quantiles",
   main="Normal Q-Q Plot")
qqline(p1.r1)

# Plot of residuals vs. fitted values
plot(p1.fitted1,p1.r1,ylab="Standardized residuals", xlab="Fitted values",
  main="Plot of residuals vs. fitted values")
abline(h=0)

# Plot of residuals vs. temperature
plot(p1.A,p1.r1,ylab="Standardized residuals", xlab="Temperature",
  main="Plot of residuals vs. temperature")
abline(h=0)

# Plot of residuals vs. pressure
plot(p1.B,p1.r1,ylab="Standardized residuals", xlab="Pressure",
  main="Plot of residuals vs. pressure")
abline(h=0)
 
#Comment: plots indicates the violation of constant-variance assumption 

#c.Use the Tukey's test to determine which levels of the pressure factor are significantly different for the data?
p1.fit2<-aov(p1.y~p1.Af+p1.Bf,
  contrasts=list(p1.Af=contr.sum,p1.Bf=contr.sum))
TukeyHSD(p1.fit2)

#at level of 0.05, all the levels of pressure are different significantly 

#d/ Under what conditions would you operate this process?
#interaction term is not significant 
p4.fit1<-lm(p1.y~p1.A+I(p1.A^2)+p1.B+I(p1.B^2))
summary(p4.fit1)
anova(p4.fit1)

press0<-seq(200,230,by=6)

press1<-rep(press0,3)

matr1<-factor(rep(1:3,each=length(press0)))

new.data<-data.frame(p1.Af=matr1,p1.B=press1)

predicted.values<-predict(p4.fit1,new.data)

interaction.plot(press1,matr1,predicted.values,type="l",pch=19,fixed=T,
   xlab="Pressure", trace.lab="Temperature",ylab="Pressure",
   main="Yield of Chemical Process")

#No interaction in interaction plot. 
#So we can operate Tukey's test to determine which levels of pressure factor 
#are significantly different for the data.
```
```{r}
#Question5
#A.Analyze the data.alpha=0.05
y2 <- c(2.70,2.78,2.45,2.49,2.60,2.72,2.75,2.86,
        2.83,2.86,2.85,2.80,2.86,2.87,2.94,2.88)
speed <- rep(c(125,200),each=8)
rate <- rep(rep(c(0.015,0.030,0.045,0.060),each=2),each=2)

speedf <-factor(speed)
ratef <- factor(rate)

y2.fit<- lm(y2~ speedf*ratef, contrasts=list(speedf=contr.sum,ratef=contr.sum)
             )
anova(y2.fit)
#I don't know why it doesn't construct the interaction of A:B in ANOVA
```



